### 2022-01-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@8da19da65b247ecf582d2d2538cadaede0352ed1 🚀

As Herbert (1965) rightfully pointed out, ``{[}y{]}ou can't draw neat
lines around planet-wide problems''; in this regard, any inference of a
metaweb at large scales must contend with several interesting and
interwoven families of problems.

The first is the spatial and taxonomic limit of the metaweb to embed and
transfer. If the initial metaweb is too narrow in scope, notably from a
taxonomic point of view, the chances of finding another area with enough
related species (through phylogenetic relatedness or similarity of
functional traits) to make a reliable inference decreases; this would
likely be indicated by large confidence intervals during estimation of
the values in the low-rank space, but the lack of well documented
metawebs is currently preventing the development of more concrete
guidelines. The question of phylogenetic relatedness and dispersal is
notably true if the metaweb is assembled in an area with mostly endemic
species, and as with every predictive algorithm, there is room for the
application of our best ecological judgement. Conversely, the metaweb
should be reliably filled, which assumes that the \(S^2\) interactions
in a pool of \(S\) species have been examined, either through literature
surveys or expert elicitation.

\textbf{TK} Supp. Mat. 1 provides some guidance as to the type of
sampling effort that should be prioritized. While RDPG was able to
maintain very high predictive power when interactions were missing, the
addition of false positive interactions was immediately detected; this
suggests that it may be appropriate to err on the side of ``too many''
interactions when constructing the initial metaweb to be transferred.

The second series of problems are related to determining which area
should be used to infer the new metaweb in, as this determines the
species pool that must be used.

\textbf{TK} In our application, we focused on the mammals of Canada. The
upside of this approach is that information at the country level is
likely to be required by policy makers and stakeholders for their
biodiversity assessment, as each country tends to set goals at the
national level (Buxton et al., 2021) for which quantitative instruments
are designed (Turak et al., 2017), with specific strategies often
enacted at smaller scales (Ray et al., 2021). And yet, we do not really
have a satisfying answer to the question of ``where does a food web
stop?''; the current most satisfying solutions involve examining the
spatial consistency of network area relationships (Fortin et al., 2021;
see \emph{e.g.} Galiana et al., 2018, 2019, 2021), which is of course
impossible in the absence of enough information about the network
itself. This suggests that an \emph{a posteriori} refinement of the
results may be required, based on a downscaling of the metaweb.

The final family of problems relates less to the availability of data or
quantitative tools, and more to the praxis of spatial ecology. Operating
under the context of national divisions, in large parts of the world,
reflects nothing more than the legacy of settler colonialism. Indeed,
the use of ecological data is not an apolitical act (Nost \& Goldstein,
2021), as data infrastructures tend to be designed to answer questions
within national boundaries, and their use both draws upon and reinforces
territorial statecraft; as per Machen \& Nost (2021), this is
particularly true when the output of ``algorithmic thinking''
(\emph{e.g.} relying on machine learning to generate knowledge) can be
re-used for governance (\emph{e.g.} enacting conservation decisions at
the national scale). We therefore recognize that methods such as we
propose operate under the framework that contributed to the ongoing
biodiversity crisis (Adam, 2014), reinforced environmental injustice
(Choudry, 2013; Domínguez \& Luoma, 2020), and on Turtle Island
especially, should be replaced by Indigenous principles of land
management (Eichhorn et al., 2019; No'kmaq et al., 2021). As we see
AI/ML being increasingly mobilized to generate knowledge that is lacking
for conservation decisions (\emph{e.g.} Lamba et al., 2019; Mosebo
Fernandes et al., 2020), our discussion of these tools need to go beyond
the technical, and into the governance consequences they can have.

>>>>

The goal of metaweb inference is to provide information about the
interactions between species at a large spatial scale. But as Herbert
(1965) rightfully pointed out, ``{[}y{]}ou can't draw neat lines around
planet-wide problems''; any inference of a metaweb at large scales must
contend with several novel, and interwoven, families of problems.

The first is the spatial and taxonomic limit of the metaweb to embed and
transfer. If the initial metaweb is too narrow in scope, notably from a
taxonomic point of view, the chances of finding another area with enough
related species (through phylogenetic relatedness or similarity of
functional traits) to make a reliable inference decreases; this would
likely be indicated by large confidence intervals during estimation of
the values in the low-rank space, or by non-overlapping trait
distributions in the known and unknown species. The lack of well
documented metawebs is currently preventing the development of more
concrete guidelines. The question of phylogenetic relatedness and
dispersal is notably true if the metaweb is assembled in an area with
mostly endemic species, and as with every predictive algorithm, there is
room for the application of our best ecological judgement.

The second series of problems are related to determining which area
should be used to infer the new metaweb in, as this determines the
species pool that must be used. Metawebs can be constructed by assigning
interactions in a list of species within geographic boundaries. The
upside of this approach is that information at the country level is
likely to be required for biodiversity assessments, as countries set
goals at the national level (Buxton et al., 2021), and as quantitative
instruments are designed to work at these scales (Turak et al., 2017);
specific strategies are often enacted at smaller scales, nested within a
specific country (Ray et al., 2021). But there is no guarantee that
these boundaries are meaningful. In fact, we do not have a satisfying
answer to the question of ``where does a food web stop?''; the most
promising solutions involve examining the spatial consistency of network
area relationships (Fortin et al., 2021; see \emph{e.g.} Galiana et al.,
2018, 2019, 2021), which is impossible in the absence of enough
information about the network itself. This suggests that infered metaweb
should be futher downscaled to allow \emph{a posteriori} analyses.

The final family of problems relates less to ecological concepts and
more to the praxis of ecological research. Operating under the context
of national divisions, in large parts of the world, reflects nothing
more than the legacy of settler colonialism. Indeed, the use of
ecological data is not an apolitical act (Nost \& Goldstein, 2021), as
data infrastructures tend to be designed to answer questions within
national boundaries, and their use both draws upon and reinforces
territorial statecraft; as per Machen \& Nost (2021), this is
particularly true when the output of ``algorithmic thinking''
(\emph{e.g.} relying on machine learning to generate knowledge) can be
re-used for governance (\emph{e.g.} enacting conservation decisions at
the national scale). We therefore recognize that predictive approaches
deployed at the continental scale, no matter their intent, originate in
the framework that contributed to the ongoing biodiversity crisis (Adam,
2014), reinforced environmental injustice (Choudry, 2013; Domínguez \&
Luoma, 2020), and on Turtle Island especially, should be replaced by
Indigenous principles of land management (Eichhorn et al., 2019; No'kmaq
et al., 2021). As we see AI/ML being increasingly mobilized to generate
knowledge that is lacking for conservation decisions (\emph{e.g.} Lamba
et al., 2019; Mosebo Fernandes et al., 2020), our discussion of these
tools need to go beyond the technical, and into the governance
consequences they can have.
